### 2022-03-31 text/article_vedrovice.Rmd
text and references before review
update texts

The Cemetery of Vedrovice is one of the earliest and most intensively studied
*Linearbandkeramik* (LBK) cemeteries.
Recent articles use evidence from the cemetery in a broader discussion
of post-marital residence patterns [@hrncir2020],
sexual inequalities and gender in the Neolithic [@masclanslatorre2020; @bickle2020].
Here we focus on an exploration of the relationship between three proposed *layers* of
identity, biological sex and age of the body determined from skeletal remains
by anthropological methods and the origin of the body based on the strontium isotope
signal.
We explore whether the structure observed in these three categories is
mirrored in the grave good associations and spatial organization within the
cemetery.
The analysis is set into a broader discussion of gender relationships and identity
in Neolithic communities.

@@

<!-- site -->
The Vedrovice site is situated in south Moravia (Czech Republic).
The earliest known finds and archaeological explorations here are from
the 19th century, but excavations were conducted in the second half of 20th century.
There are two known burial areas, *Za Dvorem* and the larger *Široká u Lesa* and
a settlement area.
For more details on the site, see @masclanslatorre2020 or the
original publication by @podborsky2002.
The material analyzed here comes from the *Široká u lesa* cemetery in Vedrovice.

<!-- sample (burials, artifacts) -->
Originally, this burial ground consisted of at least 108 burials.
Some of the burials were disturbed by later settlement and burial pits in the
area of the cemetery, amateur excavations during the 19th century and later
agricultural activities.
These burials are excluded from most of the analysis due to missing and/or
incomplete data and only a sample of 78 well-preserved LBK burials is analyzed here.
Most of the data was obtained from the original publication [@podborsky2002],
and the anthropological determination of the sex and age of the bodies was corrected based on an
analysis by @dockalova2008. Carbon, nitrogen and strontium isotope
ratios were collected in @whittle2013 and @richards2008.
Further details on the use and function of the artifacts were gathered from
@masclanslatorre2020.
Information on the localness of various bodies was adapted from @hrncir2020.

>>>>

The Cemetery of Vedrovice is one of the earliest and most intensively studied
*Linearbandkeramik* (LBK) cemeteries in the Czech Republic.
Recent articles use evidence from the cemetery in a broader discussion
of post-marital residence patterns [@hrncir2020],
sexual inequalities and gender in the Neolithic [@masclanslatorre2020; @bickle2020].
Here we focus on an exploration of the relationship between three proposed *layers* of
identity, biological sex and age of the body determined from skeletal remains
by anthropological methods and the origin of the body based on the strontium isotope
signal.
We explore whether the structure observed in these three categories is
mirrored in the grave good associations and spatial organization within the
cemetery.
The analysis is set into a broader discussion of gender relationships and identity
in Neolithic communities.

@@

<!-- site -->
The Vedrovice site is situated in south Moravia, Czech Republic.
The earliest known finds and archaeological explorations here are from
the 19th century, but excavations were conducted in the second half of 20th century.
There are two known burial areas, *Za Dvorem* and the larger *Široká u Lesa* and
a settlement area.
For more details on the site, see @masclanslatorre2020 or the
original publication by @podborsky2002.
The material analyzed here comes from the *Široká u lesa* cemetery in Vedrovice.

<!-- sample (burials, artifacts) -->
Originally, this burial ground consisted of at least 96 LBK burials and
one Lengyel culture cremation [@ondrus2002a, 21].
Some of the LBK burials were disturbed by later settlement pits in the
area of the cemetery (4 burials), amateur excavations during the 19th century
(11 burial pits with recent fills) and later agricultural activities
(at least 6 burials disturbed by deep deep plowing) or combinations of these.
These burials are excluded from the analysis of artifact combinations due to missing and/or
incomplete data and only included in the spatial analysis.
This leaves us with a sample of of 76 well-preserved LBK burials in case of
the artifact combinations (79 % of the LBK burials) and
Most of the data was obtained from the original publication [@ondrus2002a],
and the anthropological determination of the sex and age of the bodies was corrected based on an
analysis by @dockalova2008. Carbon, nitrogen and strontium isotope
ratios were collected in @whittle2013 and @richards2008.
Further details on the use and function of the artifacts were gathered from
@masclanslatorre2020.
Information on the localness of various bodies was adapted from @hrncir2020.
